COVID-19 Policy-Compliance in Europe

### 2020-12-24 Draft.Rmd
Compliance Index added, and policyMobility_data_Europe Dataframe renamed to just data. Rmd draft updated from previous version with MobilityIndex as the dependent variable

---
title: "Assignment 3 - Draft"
author: "Philipp Holz"
date: 'December 12, 2020'
output: 
  github_document:
    toc: true
    keep_html: true
bibliography: References.json

@@

```{r libraries, include=FALSE}
# Loading the libraries ----
library(dplyr)
library(ggplot2)
library(scales)
library(lme4)
```

```{r data, eval=TRUE, include=FALSE}

# Data Wrangling -----
## Adding the Policy Data -----
### Read in the Data
policy_data <- read.csv("C:/Users/holzp/Desktop/RProject/PolicyData.csv")

@@

## Adding the Mobility Data ----
### Reading in the data
mobility_data <- read.csv("C:/Users/holzp/Desktop/RProject/MobilityData.csv")

@@

## Merging the Data into policyMobility_data_Europe ----
### renaming the date columns to match
names(mobility_data_Europe)[names(mobility_data_Europe)=="date"] <- "Date"
names(mobility_data_Europe)[names(mobility_data_Europe)=="country"] <- "CountryName"
names(mobility_data_Europe)[names(mobility_data_Europe)=="avg_change"] <- "MobilityIndex"

### Only selecting the columns that are interesting
mobility_data_Europe_small <- mobility_data_Europe %>% select(Date, CountryName, MobilityIndex)
policy_data_Europe_small <- policy_data_Europe %>% select(Date, CountryName, StringencyIndex)

### Merging policy and mobility data on date
policyMobility_data_Europe = mobility_data_Europe_small %>% 
  inner_join(policy_data_Europe_small, by = c("Date", "CountryName"))

@@

policyMobility_data_Europe$ComplianceIndex <- 100 - (policyMobility_data_Europe$StringencyIndex + policyMobility_data_Europe$MobilityIndex)

@@

policyMobility_data_Europe$ComplianceIndex <- rescale(policyMobility_data_Europe$ComplianceIndex, to = c(0, 100))

@@

ggplot(data = policyMobility_data_Europe) +
  geom_line(aes(x = Date, y = StringencyIndex), color = "black", linetype = "longdash", size = 0.5) +
  geom_smooth(aes(x = Date, y = -1*MobilityIndex), method = "loess", span = 1/10, 
              se = FALSE, size = 0.5, color = "black", linetype = "solid") +
  facet_wrap(~CountryName) +
  xlab("Month") +
  scale_x_date(breaks = date_breaks("3 months"),
               labels = date_format("%b\n%Y")) +
  scale_y_reverse(name = "Policy Stringency Index (dashed line)",
                  sec.axis = sec_axis(~.*-1, name = "Mobility %-change vs Baseline (solid line)")) +
  theme(
    axis.title.x = element_text(vjust = 0.5),
    axis.title.y = element_text(color = "black", size=12, vjust = 3),
    axis.title.y.right = element_text(color = "black", size=12, vjust = 3)
  ) +
  geom_hline(yintercept = 0, color = "black", size = 0.5) +
  ggtitle("Development of Policy Stringency and Mobility Reduction in Europe")
```

@@

For this step, we can approximate policy-compliance as the difference between the level of policy-stringency and the observed reduction in mobility. The boxplot shows us that, indeed, the levels of policy compliance among European countries differ.

```{r boxplot compliance, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}
## Create a Boxplot of Compliance by Country ----
ggplot(data = policyMobility_data_Europe, aes(x = CountryName, y = ComplianceIndex)) +
  geom_boxplot() +
  xlab("Country") +
  ylab("Policy Compliance Index") +
  ggtitle("Policy Compliance in Europe") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) 
```

@@

```{r facetted plot compliance, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}
## Create a faceted plot with Compliance for European Countries ----
ggplot(policyMobility_data_Europe) +
  geom_smooth(aes(x = Date, y = ComplianceIndex), method = "loess", se = FALSE, span = 1/10, 
              color = "black", linetype = "solid", size = 0.1) +
  geom_smooth(aes(x = Date, y = ComplianceIndex), method = "lm", se = FALSE, 
              color = "black", linetype = "solid", size = 1) + 
  facet_wrap(~CountryName) +
  xlab("Month") +
  scale_x_date(breaks = date_breaks("3 months"),
               labels = date_format("%b\n%Y")) +
  scale_y_continuous(name = "Policy Compliance Index") +
  theme(
    axis.title.y = element_text(color = "black", size=12, vjust = 3),
    axis.title.x = element_text(size=12, vjust = 0.5)
    ) +
  ggtitle("Policy Compliance in Europe Across Time")
```

However, as initially stated, approximating policy-compliance as the difference between the level of policy-stringency and the observed reduction in mobility cannot be more than an approximation. When plotted across countries it shows a difference in means, and when plotted across time it shows a difference in trend.

An alternative measure of policy-compliance comes to mind, that accounts for measurement-errors and country-level differences. This alternative measure would be informed by the degree to which the mobility-reduction of a given population over time can be predicted by policy-stringency. As will be shown in more detail below, such a model is a hierarchical random-effect model.

@@

Predicted Mobility Index = β₀ + β₁ × Stringency Index + β₂ × Date + β₃ × Country + ε

>>>>

---
title: "COVID-19 Policy-Compliance in Europe"
author: "Philipp Holz"
date: 'December 24, 2020'
output: 
  github_document:
    toc: true
    keep_html: true
bibliography: References.json

@@

```{r libraries, include=FALSE}
# Loading the libraries ----
library(dplyr)
library(ggplot2)
library(scales)
library(lme4)
library(broom.mixed)
library(plotly)
library(reshape2)
library(utils)
```

```{r data, eval=TRUE, include=FALSE}
# Data Wrangling -----
## Adding the Policy Data -----
### Read in the Data
policy_data <- read.csv("PolicyData.csv")

@@

### Only selecting the columns that are interesting
policy_data_Europe_small <- policy_data_Europe %>% select(Date, CountryName, StringencyIndex)

## Adding the Mobility Data ----
### Reading in the data
mobility_data <- read.csv("MobilityData.csv")

@@

### renaming the date columns to match
names(mobility_data_Europe)[names(mobility_data_Europe)=="date"] <- "Date"
names(mobility_data_Europe)[names(mobility_data_Europe)=="country"] <- "CountryName"
names(mobility_data_Europe)[names(mobility_data_Europe)=="avg_change"] <- "MobilityIndex"

### Only selecting relevant data
mobility_data_Europe_small <- mobility_data_Europe %>% select(Date, CountryName, MobilityIndex)

## Adding the Cases Data ----

###read the Dataset sheet in
cases_data <- read.csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv")

### renaming the date and countries columns to match
names(cases_data)[names(cases_data)=="dateRep"] <- "Date"
names(cases_data)[names(cases_data)=="countriesAndTerritories"] <- "CountryName"

### Forcing Date format
cases_data$Date = as.Date(as.character(cases_data$Date), format = "%d/%m/%Y")

### create dataframe with only national level of Germany, France, UK, Spain, Italy
cases_data_Europe <- cases_data %>% filter(CountryName == "Germany" |
                                             CountryName == "Italy" |
                                             CountryName == "France" |
                                             CountryName == "Spain" |
                                             CountryName == "United_Kingdom" |
                                             CountryName == "Netherlands" |
                                             CountryName == "Belgium" |
                                             CountryName == "Poland" |
                                             CountryName == "Sweden" |
                                             CountryName == "Ireland" |
                                             CountryName == "Austria" |
                                             CountryName == "Denmark")

### Changing United_Kingdom to United Kingdom
rename <- c(United_Kingdom="United Kingdom", Germany="Germany", Italy="Italy",
            France="France", Spain="Spain", Netherlands="Netherlands", Belgium="Belgium",
            Poland="Poland", Sweden="Sweden", Ireland="Ireland", Austria="Austria",
            Denmark="Denmark")

cases_data_Europe$CountryName <- as.character(rename[cases_data_Europe$CountryName])

### Only selecting the columns that are interesting
cases_data_Europe_small <- cases_data_Europe %>% select(Date, CountryName, notification_rate_per_100000_population_14.days)

## Adding the Trends Data ----
trends_data <- read.csv("GoogleTrends.csv", sep = ";")

### Forcing correct date format
trends_data$Date = as.Date(trends_data$Date, format = "%d.%m.%Y")+1

## Renaming the Sum column to Publicity
names(trends_data)[names(trends_data)=="Sum"] <- "Publicity"

## Only selecting the columns that are interesting
trends_data <- trends_data %>% select(Date, CountryName, Publicity)

## Merging policy mobility trends and cases data on date and country ----

data = policy_data_Europe_small %>% 
  full_join(mobility_data_Europe_small, by = c("Date", "CountryName")) %>%
  full_join(cases_data_Europe_small, by = c("Date", "CountryName")) %>%
  full_join(trends_data, by = c("Date", "CountryName"))

@@

data$ComplianceIndex <- 100 - (data$StringencyIndex + data$MobilityIndex)

@@

data$ComplianceIndex <- rescale(data$ComplianceIndex, to = c(0, 100))

@@

ggplot(data = data) +
  geom_line(aes(x = Date, y = StringencyIndex), color = "black", linetype = "longdash", size = 0.5) +
  geom_smooth(aes(x = Date, y = -1*MobilityIndex), method = "loess", span = 1/10, 
              se = FALSE, size = 0.5, color = "black", linetype = "solid") +
  facet_wrap(~CountryName) +
  xlab("Month") +
  scale_x_date(breaks = date_breaks("3 months"),
               labels = date_format("%b\n%Y")) +
  scale_y_reverse(name = "Policy Stringency Index (dashed line)",
                  sec.axis = sec_axis(~.*-1, name = "Mobility %-change vs Baseline (solid line)")) +
  theme(
    axis.title.x = element_text(vjust = 0.5),
    axis.title.y = element_text(color = "black", size=12, vjust = 3),
    axis.title.y.right = element_text(color = "black", size=12, vjust = 3)
  ) +
  geom_hline(yintercept = 0, color = "black", size = 0.5) +
  ggtitle("Development of Policy Stringency and Mobility Reduction in Europe")

@@

For this analysis, policy-compliance is approximated as the difference between the level of policy-stringency and the observed reduction in mobility.

`Compliance = 100 - (Policy Stringency Index - %-change in Mobility)`

The boxplot shows us that, indeed, the levels of policy compliance among European countries differ.

```{r boxplot compliance, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}
## Create a Boxplot of Compliance by Country ----
ggplot(data = data, aes(x = CountryName, y = ComplianceIndex)) +
  geom_boxplot() +
  xlab("Country") +
  ylab("Policy Compliance Index") +
  ggtitle("Policy Compliance in Europe") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) 
```

@@

```{r facetted plot compliance, echo=FALSE, message=FALSE, warning=FALSE, fig.align="center"}
## Create a faceted plot with Compliance for European Countries ----
ggplot(data) +
  geom_smooth(aes(x = Date, y = ComplianceIndex), method = "loess", se = FALSE, span = 1/10, 
              color = "black", linetype = "solid", size = 0.1) +
  geom_smooth(aes(x = Date, y = ComplianceIndex), method = "lm", se = FALSE, 
              color = "black", linetype = "solid", size = 1) + 
  facet_wrap(~CountryName) +
  xlab("Month") +
  scale_x_date(breaks = date_breaks("3 months"),
               labels = date_format("%b\n%Y")) +
  scale_y_continuous(name = "Policy Compliance Index") +
  theme(
    axis.title.y = element_text(color = "black", size=12, vjust = 3),
    axis.title.x = element_text(size=12, vjust = 0.5)
    ) +
  ggtitle("Policy Compliance in Europe Across Time")
```

@@

`Compliance Index = β₀ + β₁ × Date + β₂ × Country + β₃ × Publicity + β₄ × Cases per 100.000 per 14 days + ε`

THE FOLLOWING MUST BE UPDATED
